% $Id: CoralDB-Performance.tex,v 1.3 2007-07-31 15:53:56 jb Exp $

\subsection{Performance}
\label{performance}

% Main author: Juerg

Preliminary performance studies of the CoralDB connectivity database have been performed in a standalone
environment using both local SQLite files and the Oracle test server {\tt devdb10}. While such studies
can give an impression of the best-case performance, the performance figures that are really relevant for
ATLAS can only be determined in a test environment that includes at least the whole inner detector, if not
all of ATLAS. In particular, in order to get more realistic performance estimates when using Oracle, the number of clients
accessing the Oracle server should be similar to what is expected in the real experiment for example at the
start of a new run.

The performance numbers given in table~\ref{PERFORMANCE-table} below have been obtained by running a single instance
of the CoralDB code on the {\tt atlpix01} PC at CERN. {\tt atlpix01} is a more or less standard Intel Pentium 4, 3GHz machine with 0.5 GB of RAM. Timing measurements were performed using either the shell command {\tt time} (for long operations) or
{\tt gettimeofday()} from {\tt sys/time.h} (whose nominal resolution is 1$\mu$s). Measurements for SQLite files were
made using a file on a local disk. Measurements for Oracle were made using the {\tt devdb10} development server (the online server is expected to be significantly faster, but the actual performance will of course depend on the server's load).

Since during these studies the PC, network connections and Oracle server were shared with other users, the test results showed a large variation depending on other activities going on in parallel. Therefore the results quoted in
table~\ref{PERFORMANCE-table} are indicative at best and all numbers have large errors (perhaps at the level of 50\%). 


\begin{table}[htb]
\begin{center}
\begin{tabular}{|p{11cm}|c|c|}
\hline
What & \multicolumn{2}{|c|}{Typical execution time} \\
& SQLite & Oracle \\ \hline
\hline
Start database connection/transaction & 60ms & $<1$s \\ \hline
Find a connection                     & 0.3ms  & 10ms  \\ \hline
Traverse connectivity using PixDbCoralDB, after caching & 12$\mu$s & 12$\mu$s \\ \hline
Add or update a link using PixDbCoralDB & 0.2ms & 9ms \\ \hline
Operations on CLOBs (up to 4kB) & \multicolumn{2}{|c|}{O(tens of ms)} \\ \hline
\hline
Create new database tables and upload pit connectivity (10k connections, 7k aliases) & 10s & 4min \\ \hline
\end{tabular}
\caption[CoralDB performance summary]
{Summary of CoralDB performance for local SQLite files and for the devdb10 Oracle server.}
\label{PERFORMANCE-table}
\end{center}
\end{table}
